How to improve your relationship with your future self.

### 2016-06-06 workflow.Rmd
Version ready for Maarten

---
title: How to improve your relationship with your future self.
author:
- name: Jake Bowers
  affiliation: <jwbowers@illinois.edu>, University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team. Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic. The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
- name: Maarten Voors
  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics
bibliography: references.bib
csl: american-political-science-review.csl
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    pdf_document:
      number_sections: true
      fig_width: 7
      fig_height: 7
      fig_caption: true
      template: bowersarticletemplate.latex
      pandoc_args: [
      "--filter", "/usr/local/bin/pandoc-crossref"
    ]
...

@@

row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
good.df<-na.omit(dem.nations.df[,c("Gini2004","WVS5_protest","Nation","meanpr")])
good.df$gini04<-good.df$Gini2004/100
good.df$protest05<-good.df$WVS5_protest
```

@@

Comments --- unexecuted text inside of a script --- are a message to
collaborators (including your future self) and other consumers of your
work. In the above code chunk, I used comments to explain the lines to
readers unfamiliar with R and to remember that I had tried a different
specification but decided not to use it because adding the squared
term did not really change the substantive story arising from the
model.^[R considers text marked with `#` as a comment.]
Messages left for your future self (or near-future others) help
retrace and justify your decisions as the work moves from seminar
paper to conference paper to poster back to paper to dissertation and
onwards.

@@

For example, in [@sec:data-analys-comp], I suggested that we know
where 'that plot on page 10 'comes from by making sure we had a
`fig1.pdf` file produced from a clearly commented plain text file
called something like `fig1.R`.  An even easier solution would be to
directly include a chunk of code to produce the figure inside of the
paper itself. This paper, for example, was written in plain text using
\LaTeX~ markup with R code chunks to make things like
Figure~\ref{fig:giniprot}. This combination of LaTeX and R is called
Sweave [@leis:2005].^[Support for Sweave is included with R.]

```
This paper, for example, was written in plain text using Markdown
markup with R code chunks to make things like
Figure~\ref{fig:giniprot}. This combination of Markdown and R is
called [R markdown](http://rmarkdown.rstudio.com/).\footnote{Support for R markdown is
  included with R. For nicer documents, use the `rmarkdown` R package or click `Knit to PDF` or `Knit to HTML` using the free and open-source [RStudio IDE](http://rstudio.com) for R.}
\begin{figure}[h]
\centering
\`\`\`{r fig1plota,fig=TRUE}
##  Make a scatterplot of Protest by Inequality
with(good.df,plot(gini04,protest05,xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Mean Protest Activities\n(World Values Survey 2005--2010)'))
##  Label a few interesting points
with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
     text(gini04,protest05,labels=Nation))
\`\`\`
\caption{Protest activity by income inequality [@norris2009data].}
\label{fig:giniprot}
\end{figure}
```

```{r fig1code,echo=FALSE,fig=FALSE}
##  Make a scatterplot of Protest by Inequality
with(good.df,plot(gini04,protest05,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
                  cex=.8))

##  Label a few interesting points
with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
```

\begin{figure}[H]
  \begin{center}
```{r fig1plot, ref.label="fig1code",echo=FALSE,fig=TRUE,width=4,height=4}
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
```
    \caption{Protest activity by income inequality @norris2009data.}
    \label{fig:giniprot}
  \end{center}
\end{figure}

By using `\label{fig:giniprot}`, I do not need to keep track of
the figure number, nor do extra work when I reorganize the document in
response to reviewer suggestions. Nor do I need a separate
`fig1.R`file or `fig1.pdf` file.  Tables and other
numerical results are also possible to generate within the source
code of a scholarly paper. Those who view the code for this essay will
see how Table~\ref{tab:protest} was also generated directly from a
regression object.\footnote{@beck2010reg inspired this
  particular presentation of a linear model.}

@@

library(xtable)
thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, adjusting for average political
  rights as measured by Freedom House 1980--2000. Data from [@norris2009data].",label="tab:protest")
```

```{r printtable,echo=FALSE,results='asis'}
  print(thextab,include.colnames=FALSE,
        hline.after=NULL,
        table.placement="H",
        add.to.row=list(
          pos=list(0,3),
          command=c(
            "\\toprule \n &
             \\multicolumn{1}{c}{Coef} &
             \\multicolumn{1}{c}{Std. Err.} &
             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
            paste("& \\multicolumn{4}{c}{n: ",n,",",
                  "   resid.sd: ",signif(est.sigma),
                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
            )
          ))
```

@@

** Step 2 **

We analyze data in order to explain something about the world to other
scholars and policy makers. If we focus on explaining how we got our
computers to do data analysis to human beings, we will do a better job
with the data analysis itself: we will learn as we focus on teaching,
and we will avoid errors and save time as we ensure that others
(including our future selves) can retrace our steps. A document than
can be 'run 'to reproduce all of the analyses also instills
confidence in readers and can more effectively spur discussion and
learning and cumulation of research.

@@

@@

The principle of modularity suggests that you separate data cleaning,
processing, recoding, and merging from analysis in different files
[@nagler1995coding]. So, perhaps your analysis oriented files
will `load("cleandata.rda")`and a comment in the code will alert
the future you (among others) that `cleandata.rda`was created
from `create-cleandata.R` which in turn begins with
`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such
a data processing file will typically end with something like
`save("cleandata.rda")`so that we are doubly certain about the
provenance of the data.\footnote{Of course, if you need math or
  paragraphs to explain what is happening in these files, you might
  prefer to make them into Sweave files, for which the conventional
  extension is .Rnw. So you'd have `create-cleandata.Rnw` which
  might explain and explore the different coding decisions you made,
  perhaps involving a factor analysis and diagnostic plots.}

@@

** Step 3** We should know where the data came
from and what operations were performed on which set of data.

In the good old days, when we executed our LISREL code in batch mode,
we had no choice but to tell the machine clearly, in a few easy to
understand and informative lines, what files (with filenames no longer
than 8 characters) to use:

@@

# Version control prevents clobbering and reconciles history. {#sec:vers-contr-prev}

Group work requires version control.^[@fredrickson2011tpm and
@healy2011tpm in this issue also explain what version control is and why we
might want to use it.] Many people are familiar with the track changes
feature in modern WYSIWYG word processors or the fact that Dropbox allows one
to recover previous versions of files. These are both kinds of version control.
More generally, when we collaborate, we'd like to do a variety of actions with
our shared files. Collaboration on data analytic projects is more productive
and better when (1) it is easy to see what has changed between versions of
files; (2) members of the team feel free to experiment and then to dump parts
of the experiment in favor of previous work while merging the successful parts
of the experiment into the main body of the paper; (3) the team can produce
have 'releases 'of the same document (one to MPSA, one to APSR, one to your
parents) without spawning many possibly conflicting copies of the same
document; (4) people can work on the same files at the same time without
conflicting with one another (and can reconcile their changes without too much
confusion and clobbering).  Clobbering is what happens when your future self or
your current collaborator saves an old version of a file over a new version,
erasing good work by accident.

Of course if you rely on Dropbox or 'track changes 'for version
control, you must communicate with other folks in your group before
you edit existing files. Only one of you can edit and save a given
file at a time. This prevents your work (or your colleagues work) from
getting lost when you both try to save the same file on top of each
other. If you find that you need to work on the same files at the same
time, then you should work on establishing your own shared version
control system. Free options include launchpad, github, sourceforge
for open source projects (i.e. papers you are writing which you are
happy to share with others as you write). Each of those services
include paid versions too. One may also use Dropbox as a kind of
server for version control: for example, one may copy files from the
Dropbox directory into a local working directory so as to avoid
clobbering and then work on merging changes by hand before copying
back to the Dropbox directory and replacing existing files.

We use git with our own research group, and I use it for all of
my own projects.MORE ON GITHUB

@@

** Step 5** Writing is rewriting. Thus, all writing
involves versions. When we collaborate with ourselves and others we
want to avoid clobbering and we want to enable graceful reconciliation
of rewriting. One can do these things with formal systems of software
(like subversion or git or bazaar) or with formal systems of file
naming, file comparing, and communication or, even better, with
both. In either case, plain text files will make such tasks easier,
will take up less disk space, and be easier to read for the future
you.

@@

Now, back to that famous article of 2018. After reading the conference
paper critique of 2021 (that came from the seminar paper of 2020), the
statisticians at the UN worry about the bootstrap confidence intervals
presented in the original paper.^[Perhaps they should be
  worried about the deeper substantive critiques offered by the
  student, but they are statisticians and so focus on the stats. The
  policy makers of 2021 were cowed by the methodological virtuosity of
  the 2018 article, and so, even though they had the same substantive
  concerns as the student, they kept their mouths shut at the
  mini-conference to avoid looking dumb in front of their bosses.] So,
now the authors would like to evaluate their bootstrap procedure. Although
nice code exists for bootstrapping linear models, no nice code exists
to bootstrap the bootstrap. Of course, the code required is not
complex, but since they are writing custom code they worry about
getting it right. As they've struggled to respond to the critiques of
their paper, they've had lots of time to appreciate problems arising
from bugs, errors, and typos in data analysis and code.

Now, if they had a moment to think in between teaching that new class,
reading books for an awards committee, evaluating application files
for the admissions committee, feeding popsicles to a sick child, and
undertaking the odd bit of research, they might say to themselves,
"Before I write new code, I should write a test of the code. I should
write a little bit of code that lets me know that my double-bootstrap
procedure actually does what it is supposed to do."

Of course, this idea, like most others, is not new. The desire to
avoid error looms large when large groups of programmers write code
for multi-million dollar programs. The idea of
[test driven development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea that one ought to create tests of
[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns. For the social scientist
collaborating with her future self and/or a small group of
collaborators, here is an example of this idea in a very simple form:
Say I want to write a function to multiply a number by 2. If my
function works, when I give it the number 4, I should see it return
the number 8 and when I give it -4, I should get -8.

@@

@@

I have experimented with three systems so far: (1) for one paper we simply
included a Sweave document and data files into a compressed archive
[@bowers2005dataverse]; (2) for another more computing intensive paper we
assembled a set of files that enabled reproduction of our results using the
`make` system [@bowers2008dataverse]; and (3) recently I have tried the
"compendium" approach [@gentleman2005reproducible,@gentleman2007statistical]
which embeds an academic paper within the R package system
[@bowers2011dataverse]. The benefit of this last approach is that one is not
required to have access to a command line for `make`: the compendium is
downloadable from within R using `install.packages()` and is viewable using the
`vignette()` function in any operating system than runs R.^[Notice that my
reproduction archives and/or instructions for using them are hosted on the
[Dataverse](http://thedata.org/book/learn-about-project), which is another
system designed to enhance academic collaboration across time and space.]  The
idea that one ought to be able to install and run and use an academic paper
just as one installs and uses statistical software packages is very attractive,
and I anticipate that it will become ever easier to turn papers into R packages
as creative and energetic folks turn their attention to the question of
reproducible research.

@@

But what if the UN never hears of your work, or, by some cruel fate, your
article does not spawn debate? Why then would you spend time to
communicate with your future self and others? My own answer to this
question is that I want my work to be credible and useful to myself
and other scholars even if each article does not immediately change
the world.  What I report in my data analyses should have two main
characteristics: (1) the findings of the work should not be a matter
of opinion; and (2) other people should be able to reproduce the findings. That is,
the work represents a shared
experience --- and an experience shared without respect to the
identities of others (although requiring some common technical
training and research resources).

% Such work should make us change how
% we act --- or at least, it ought to stand on stronger epistemological
% ground than other claims about experiences which ought to be shared or
% shareable.

Assume we want others to believe us when we say something. More
narrowly, assume we want other people to believe us when we say
something about data: 'data 'here can be words, numbers, musical
notes, images, ideas, etc \ldots The point is that we are making some
claims about patterns in some collection of stuff. Now, it might be
easy to convince others that 'this collection of stuff is different
from that collection of stuff 'if those people were looking over our
shoulders the whole time that we made decisions about collecting the
stuff and broke it up into understandable parts and reorganized and
summarized it. Unfortunately, we can't assume that people are willing
to shadow a researcher throughout her career. Rather, we do our work
alone or in small groups and want to convince other distant and future
people about our analyses.

@@

In the end, following these practices and those recommended by
@fredrickson2011tpm and @healy2011tpm in this issue
allows your computerized analyses of your collections of stuff to be
credible.  Finally, if the UN quibbles with your analyses, your future
self can shoot them the archive required to reproduce your
work.^[Since you used plain text, the files will still be
  intelligible, analyzed using commented code so that folks can
  translate to whatever system succeeds R, or since you used R, you
  can include a copy of R and all of the R packages you used in your
  final analyses in 2018 in the archive itself. You can even throw in
  a copy of Ubuntu 10 and an open source virtual machine running the
  whole environment.] You can say, 'Here is everything you need to
reproduce my work.' To be extra helpful you can add 'Read the README
file for further instructions.' And then you can get on with your life:
maybe the next great idea will occur when your 4-year-old asks a wacky
question after stripping and painting her overly cooperative
1-year-old brother purple, or teaching a class, or in a coffee shop,
or on a quiet walk.

>>>>

---
title: How to improve your relationship with your future self.
author:
- name: Jake Bowers
  affiliation: <jwbowers@illinois.edu>, University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team. Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic. The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
- name: Maarten Voors
  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics
bibliography: references.bib
csl: american-political-science-review.csl
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    pdf_document:
      keep_tex: true
      number_sections: true
      fig_width: 7
      fig_height: 7
      fig_caption: true
      template: bowersarticletemplate.latex
      pandoc_args: [
      "-M", "secPrefix=\\S",
      "--filter", "/usr/local/bin/pandoc-crossref"
      ]
      md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
...

@@

row.names(dem.nations.df)<-dem.nations.df$Nation ##make the row names more helpful
good.df<-na.omit(dem.nations.df[,c("Gini2004","WVS5_protest","Nation","meanpr")])
good.df$gini04<-good.df$Gini2004/100
good.df$protest05<-good.df$WVS5_protest
```

@@

Comments, unexecuted text inside of a script, are a message to
collaborators (including your future self) and other consumers of your
work. In the above code chunk, I used comments to explain the lines to
readers unfamiliar with R and to remember that I had tried a different
specification but decided not to use it because adding the squared
term did not really change the substantive story arising from the
model.^[R considers text marked with `#` as a comment.]
Messages left for your future self (or near-future others) help
retrace and justify your decisions as the work moves from seminar
paper to conference paper to poster back to paper to dissertation and
onwards.

@@

For example, in [@sec:data-analys-comp], I suggested that we know
where 'that plot on page 10 'comes from by making sure we had a
`fig1.pdf` file produced from a clearly commented plain text file
called something like `fig1.R`.  An even easier solution would be to
directly include a chunk of code to produce the figure inside of the
paper itself. This paper, for example, was written in plain text using
markdown markup with R code chunks to make things like
Figure \ref{fig:giniprot}. This combination of markdown and R is called
[R Markdown](http://rmarkdown.rstudio.org).

```{r fig1plotcode,eval=FALSE,echo=TRUE}
##  Make a scatterplot of Protest by Inequality
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
with(good.df,plot(gini04,protest05,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
                  cex=.8))
##  Label a few interesting points
with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
```

```{r giniprot,message=FALSE,echo=FALSE,fig=TRUE,include=TRUE,width=4,height=4,fig.cap="Average number of Protest activity by income inequality.",out.width=".5\\textwidth",fig.env="figure",fig.lp="fig:",fig.pos="H",fig.align="center"}
##  Make a scatterplot of Protest by Inequality
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
with(good.df,plot(gini04,protest05,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
                  cex=.8))
##  Label a few interesting points
with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
```

Markdown (and LaTeX and HTML) all have ways to cross-reference within a
document. For example, by using `\label{fig:giniprot}` in LaTeX or
`{#fig:giniprot}` in the pandoc version of markdown, I do not need to keep
track of the figure number, nor do extra work when I reorganize the document in
response to reviewer suggestions. Nor do I need a separate `fig1.R`file or
`fig1.pdf` file.  Tables and other numerical results are also possible to
generate within the source code of a scholarly paper. Those who view the code
for this essay will see how Table \ref{tab:protest} was also generated directly
from a regression object.\footnote{@beck2010reg inspired this particular
presentation of a linear model.}

@@

library(xtable)
thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, adjusting for average political
  rights as measured by Freedom House 1980--2000. Data from [@norris2015data].",label="tab:protest")
```

```{r printtable,echo=FALSE,results='asis',message=FALSE}
  print(thextab,include.colnames=FALSE,
        hline.after=NULL,
        table.placement="H",
	comment=FALSE,
        add.to.row=list(
          pos=list(0,3),
          command=c(
            "\\toprule \n &
             \\multicolumn{1}{c}{Coef} &
             \\multicolumn{1}{c}{Std. Err.} &
             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
            paste("& \\multicolumn{4}{c}{n: ",n,",",
                  "   resid.sd: ",signif(est.sigma),
                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
            )
          ))
```

@@

**Step 2**
We analyze data in order to explain something about the world to other
scholars and policy makers. If we focus on explaining how we got our
computers to do data analysis to human beings, we will do a better job
with the data analysis itself: we will learn as we focus on teaching,
and we will avoid errors and save time as we ensure that others
(including our future selves) can retrace our steps. A document than
can be 'run' to reproduce all of the analyses also instills
confidence in readers and can more effectively spur discussion and
learning and cumulation of research.

@@

[MAARTEN IDEAS HERE? PLUS ANOTHER SECTION ABOUT ORGANIZATION?]

@@

The principle of modularity suggests that you separate data cleaning,
processing, recoding, and merging from analysis in different files
[@nagler1995coding]. So, perhaps your analysis oriented files will
`load("cleandata.rda")`and a comment in the code will alert the future you
(among others) that `cleandata.rda`was created from `create-cleandata.R` which
in turn begins with
`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such a data
processing file will typically end with something like
`save("cleandata.rda")`so that we are doubly certain about the provenance of
the data.\footnote{Of course, if you need math or paragraphs to explain what is
happening in these files, you might prefer to make them into R markdown or
Sweave files, for which the conventional extension is .Rmd or .Rnw
respectively. So you'd have `create-cleandata.Rnw` written as a mixture of
laTeX and R which might explain and explore the different coding decisions you
made, perhaps involving a factor analysis and diagnostic plots.}

@@

**Step 3** We should know where the data came
from and what operations were performed on which set of data.

In the good old days, when we executed our LISREL code in batch mode,
we had no choice but to tell the machine clearly, in a few easy to
understand and informative lines, what files (with filenames no longer
than 8 characters) to use. See for example, this extremely clear code snippet:

@@

# Version control prevents clobbering, reconciles history, and helps organize work. {#sec:vers-contr-prev}

Group work requires version control.^[@fredrickson2011tpm and
@healy2011tpm in this issue also explain what version control is and why we
might want to use it.] Many people are familiar with the track changes
feature in modern WYSIWYG word processors or the fact that Dropbox allows one
to recover previous versions of files. These are both kinds of version control.
More generally, when we collaborate, we'd like to do a variety of actions with
our shared files. Collaboration on data analytic projects is more productive
and better when (1) it is easy to see what has changed between versions of
files; (2) members of the team feel free to experiment and then to dump parts
of the experiment in favor of previous work while merging the successful parts
of the experiment into the main body of the paper; (3) the team can produce
have 'releases' of the same document (one to MPSA, one to APSR, one to your
parents) without spawning many possibly conflicting copies of the same
document; (4) people can work on the same files at the same time without
conflicting with one another (and can reconcile their changes without too much
confusion and clobbering).  Clobbering is what happens when your future self or
your current collaborator saves an old version of a file over a new version,
erasing good work by accident.

Of course if you rely on Dropbox or 'track changes' in Word for version
control, you must communicate with other folks in your group before you edit
existing files. Only one of you can edit and save a given file at a time. This
prevents your work (or your colleagues work) from getting lost when you both
try to save the same file on top of each other. If you find that you need to
work on the same files at the same time, then you should work on establishing
your own shared version control system. Free options include launchpad, github,
sourceforge for open source projects (i.e. papers you are writing which you are
happy to share with others as you write). As of this writing, Google Docs does
an excellent job with shared editing of files as long as your are online. Each
of those services include paid versions too. One may also use Dropbox as a kind
of server for version control: for example, one may copy files from the Dropbox
directory into a local working directory so as to avoid clobbering and then
work on merging changes by hand before copying back to the Dropbox directory
and replacing existing files. MORE PERHAPS ON GOOGLE? LESS ON DROPBOX?

We use git with our own research group, and I use it for all of
my own projects. MORE ON GITHUB

@@

**Step 5** Writing is rewriting. Thus, all writing
involves versions. When we collaborate with ourselves and others we
want to avoid clobbering and we want to enable graceful reconciliation
of rewriting. One can do these things with formal systems of software
(like subversion or git or bazaar) or with formal systems of file
naming, file comparing, and communication or, even better, with
both. In either case, plain text files will make such tasks easier,
will take up less disk space, and be easier to read for the future
you.

@@

Anyone writing custom code should worry about getting it right. The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis
and code.

Now, if one has a moment to think in between teaching that new class,
reading books for an awards committee, evaluating application files
for the admissions committee, feeding popsicles to a sick child, and
undertaking the odd bit of research, one might say to oneself,
"Before I write new code, I should write a test of the code. I should
write a little bit of code that lets me know that my double-bootstrap
procedure actually does what it is supposed to do."

Of course, this idea, like most others, is not new. The desire to
avoid error looms large when large groups of programmers write code
for multi-million dollar programs. The idea of
[test driven development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea that one ought to create tests of
[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns.^[There are now R packages to help R package writers do this, see for example <https://github.com/hadley/testthat> and the article on it <https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.] For the social scientist
collaborating with her future self and/or a small group of
collaborators, here is an example of this idea in a very simple form:
Say I want to write a function to multiply a number by 2. If my
function works, when I give it the number 4, I should see it return
the number 8 and when I give it -4, I should get -8.

@@

MAYBE SOMETHING HERE ON `stopifnot` and recoding data.

@@

I have experimented with three systems so far: (1) for one paper we simply
included a Sweave document and data files into a compressed archive
[@bowers2005dataverse]; (2) for another more computing intensive paper we
assembled a set of files that enabled reproduction of our results using the
`make` system [@bowers2008dataverse]; (3) I have tried the "compendium"
approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
academic paper within the R package system [@bowers2011dataverse]; and (4)
recently I have started organizing my work on github which enables me to make
"releases" of the project which should enable smooth reproduction of all of the
products of the research (simulations, data analyses, tables, figures, etc..).
The benefit of the compendium approach is that one is not required to have
access to a command line for `make`: the compendium is downloadable from within
R using `install.packages()` and is viewable using the `vignette()` function in
any operating system than runs R.^[Notice that my reproduction archives and/or
instructions for using them are hosted on the
[Dataverse](http://thedata.org/book/learn-about-project), which is another
system designed to enhance academic collaboration across time and space.]  The
idea that one ought to be able to install and run and use an academic paper
just as one installs and uses statistical software packages is very attractive,
and I anticipate that it will become ever easier to turn papers into R packages
as creative and energetic folks turn their attention to the question of
reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]

@@

But what if no one ever hears of your work, or, by some cruel fate, your
article does not spawn debate? Why then would you spend time to
communicate with your future self and others? My own answer to this
question is that I want my work to be credible and useful to myself
and other scholars even if each article does not immediately change
the world.  What I report in my data analyses should have two main
characteristics: (1) the findings of the work should not be a matter
of opinion; and (2) other people should be able to reproduce the findings. That is,
the work represents a shared
experience --- and an experience shared without respect to the
identities of others (although requiring some common technical
training and research resources).

Assume we want others to believe us when we say something. More narrowly,
assume we want other people to believe us when we say something about data:
'data' here can be words, numbers, musical notes, images, ideas, etc \ldots The
point is that we are making some claims about patterns in some collection of
stuff. For example, when I was invited into the homes and offices of ordinary
people in Chile in 1991, the "stuff" was recordings of conversations that we
had about life during the first year of democracy and I was comparing the
responses of people who had experienced the Pinochet dictatorship differently.
Now, it might be easy to convince others that 'this collection of stuff is
different from that collection of stuff' if those people were looking over our
shoulders the whole time that we made decisions about collecting the stuff and
broke it up into understandable parts and reorganized and summarized it.
Unfortunately, we can't assume that people are willing to shadow a researcher
throughout her career. Rather, we do our work alone or in small groups and want
to convince other distant and future people about our analyses.

@@

In the end, following these practices and those recommended by
@fredrickson2011tpm and @healy2011tpm among others working on these topics
allows your computerized analyses of your collections of stuff to be
credible.  Finally, if the someone quibbles with your analyses, your future
self can shoot them the archive required to reproduce your
work.^[Since you used plain text, the files will still be
  intelligible, analyzed using commented code so that folks can
  translate to whatever system succeeds R, or since you used R, you
  can include a copy of R and all of the R packages you used in your
  final analyses in the archive itself. You can even throw in
  a copy of whatever version of linux you used and an open source virtual machine running the
  whole environment.] You can say, 'Here is everything you need to
reproduce my work.' To be extra helpful you can add 'Read the README
file for further instructions.' And then you can get on with your life:
maybe the next great idea will occur when your 4-year-old asks a wacky
question after stripping and painting her overly cooperative
1-year-old brother purple, or teaching a class, or in a coffee shop,
or on a quiet walk.
